\documentclass[11pt]{article}
\linespread{1.5}
\usepackage{fullpage}
\usepackage{amsmath,theorem,amssymb,graphicx, pgfplots, tabularx, placeins}
\usepackage[semicolon,authoryear]{natbib}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{csquotes}
\usepackage{epstopdf}

\newcommand{\bit}{\begin{itemize}}
\newcommand{\eit}{\end{itemize}}

\usepackage{bibentry}

\newcommand{\lb}{\label}
\newtheorem{thm}{Theorem}
\newtheorem{prop}{Proposition}
\newtheorem{definition}{Definition}

\title{Preliminary Detalied Syllabus for Macro II Part II, Spring 2026}
\author{Erik \"{O}berg}
\date{}

\begin{document}
\bibliographystyle{apalike}
\nobibliography{bibtex}
\maketitle

\section*{Overview}

\begin{itemize}
	\item The course introduces workhorse models used in the i) business cycle, ii) frictional labor market and iii) incomplete markets literatures. 
	Emphasis will be put on developing analytical techniques for actively working with these models as well as understanding their core implications. 
	We will also repeatedly explore the usage of micro data to inform macroeconomic models. 
	
	\item 14 lectures: 
	\bit
		\item 6 lectures (20 hours) on business cycle models
		\item 4 lectures (13 hours) on frictional labor markets
		\item 3 lectures (7 hours) on incomplete asset markets
		\item 1 summing-up/Q\&A session
	
	\eit
	
	\item 7 TA sessions: 1 tutorial session on using Dynare / 6 sessions on graded problem sets
	
	\item Basic proficiency with Matlab is assumed; there are a lot of tutorials you can check online. 
	
	\item The Ph.D. Macro Book (December 2025 version) is our main textbook. Core readings are marked with a *. The other listed readings are highly recommended. In addition to this material, we will discuss some additional papers in class, these readings are for the interested.
\end{itemize}


\section*{Lecture 1 - RBC: Business Cycle Facts and the baseline RBC model (3 hours)}
\begin{enumerate}
\item Business cycle facts

\item Math preliminaries

\item The Real Business Cycle model: Setup and solution

\item The Real Business Cycle model: Analysis
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 14 (including the appendix).
	\item Eric Sims's lecture notes entitled ``Preliminaries'', ``Log-lineariziation'', ``Using Dynare'', ``RBC (qualitative)'' and ``RBC (quantitative)''
	\item \bibentry{King1999}
	\item \bibentry{Kydland1982}
	\item \bibentry{King1984}
\end{itemize}


\section*{Lecture 2 -  RBC: accounting, measurement, extensions}
\begin{enumerate}
\item Business cycle accounting

\item Measuring technology shocks

\item Labor market extensions
\bit
\setlength\itemsep{0.5em} 

\item Employment lotteries

\item GHH preferences

\eit
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 14 (including the appendix).
	\item *\bibentry{Chari2007}
	\item \bibentry{King1999}
	\item \bibentry{Basu2006}
	\item \bibentry{Hansen1985}
	\item \bibentry{Rogerson1988}
	\item \bibentry{greenwood1988investment}
\end{itemize}


\section*{Lecture 3 - RBC: investment Dynamics}
\begin{enumerate}	
	\item RBC setup with firm ownership of capital
	
	\item Neoclassical theory vs. Q theory of investment
	\bit
	\item Investment adjustment costs in the RBC model
	\eit
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 14 (including the appendix).
	\item *\bibentry{Hayashi1982}
	\item *Eric Sims's lecture notes entitled ``RBC extensions''
\end{itemize}

\section*{Lecture 4 - The New-Keynesian model: basics (3 hours)}
\begin{enumerate}
	\item Evidence concering the effects of monetary policy
	
	\item The vanilla NK model: setup and equilibrium
	
	\item Determinacy and Taylor rules
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 18 (including the appendix).
	\item *\bibentry{Ramey2016}
	\item \bibentry{Cochrane2011}
	\item \bibentry{Clarida1999} 
\end{itemize}


\section*{Lecture 5 - The New-Keynesian model: mechanics}
\begin{enumerate}
	\item Monetary policy shocks
	
	\item TFP shocks
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 18 (including the appendix).
	\item \bibentry{Clarida1999}
	\item \bibentry{Christiano2005}
	\item \bibentry{Smets2007}
\end{itemize}

\section*{Lecture 6 - The New-Keynesian model: policy}
\begin{enumerate}	
	\item Sources of ineffiency 
	
	\item Optimal monetary policy
	
	\item Quantitative NK models: A helicopter view
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 18 (including the appendix).
	\item \bibentry{Clarida1999}
	\item \bibentry{Christiano2005}
	\item \bibentry{Smets2007}
\end{itemize}




\section*{Lecture 7 - Frictional labor markets: basics (3 hours)}
\begin{enumerate}
	\item Labor markets facts: stocks, flows and prices
	
	\item Mathematical preliminaries
	
	\item Search in partial equilibrium (the McCall model)
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *\bibentry{Rogerson2005}
	\item *\bibentry{Hornstein2011}
	\item \bibentry{Rogerson2011}
	\item \bibentry{Elsby2013}
\end{itemize}


\section*{Lecture 8 - Job ladders and wage dispersion}
\begin{enumerate}
	\item The AKM regression
	\item The Burdett-Mortensen model
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *\bibentry{Rogerson2005}
	\item *\bibentry{Burdett1998}
	\item *\bibentry{Hornstein2011}
	\item \bibentry{Abowd1999}
\end{itemize}

\section*{Lecture 9 - Unemployment: statics}
\begin{enumerate}
	\item The matching function
	\item Diamond-Mortensen-Pissarides: setup and solution 
	\item Diamond-Mortensen-Pissarides: static analysis
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *\bibentry{Rogerson2005}
	\item *Ph.D. Macro Book, Chapter 20 (including the appendix).
\end{itemize}


\section*{Lecture 10 - Unemployment: efficiency and dynamics}
\begin{enumerate}
	\item Diamond-Mortensen-Pissarides: Efficiency
	
	\item Diamond-Mortensen-Pissarides: dynamic analysis
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 20 (including the appendix).
	\item *\bibentry{Rogerson2005}
	\item *\bibentry{Shimer2005}
	\item *\bibentry{Hagedorn2008}
	\item \bibentry{Hosios1990}
\end{itemize}


\section*{Lecture 11 - Incomplete markets: Basics}
\begin{enumerate}
	\item Basic facts about the distribution of income, wealth and consumption
	\item Aggregation with complete and incomplete markets
	\item Consumption-savings dynamics with incomplete markets
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 5 and 11 (including the appendix).
	\item \bibentry{Heathcote2010}
	\item \bibentry{Guvenen2014}
	\item \bibentry{Guvenen2016}
	\item \bibentry{Kuhn2020}
\end{itemize}


\section*{Lecture 12 - Incomplete markets: Buffer-stock savings (3 hours)}
\begin{enumerate}
	\item The buffer-stock savings model
	\bit
	\item Setup, computation and calibration
	\item The consumption function
	\eit	
	\item Applying the Buffer-Stock Savings Model
	\bit
	\item Gourinchas-Parker: life-cycle savings dynamics
	\item Blundell-Pistaferri-Preston: consumption and Income inequality
	\item Kaplan-Violante: illiquid assets and the marginal propensity to consume
	\eit
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 11.
	\item *\bibentry{Carroll1997}
	\item *\bibentry{Gourinchas2002}
	\item *\bibentry{blundell2008}
	\item *\bibentry{Kaplan2014}
	\item \bibentry{Deaton1991}
\end{itemize}

\section*{Lecture 13 - Incomplete markets: General equilibrium}
\begin{enumerate}
	\item The Ayiagari model
	\bit
	\item Recursive competitive equilibrium
	\item Asset convergence
	\item The Ayiagari diagram
	\eit
	\item Applications
	\bit
		\item Explaining wealth inequality
		
		\item Taxation
	\eit
\end{enumerate}

\bf Readings \normalfont
\begin{itemize}
	\item *Ph.D. Macro Book, Chapter 11.
	\item *\bibentry{Aiyagari1994}
	\item \bibentry{Huggett1993}
\end{itemize}


\section*{Lecture 14 - Course summary and pre-exam Q\&A}


%\section*{Non-required readings cited in class (roughly in order of appearance)}
%\begin{itemize}
%	\item \bibentry{Shimer2012}
%	\item \bibentry{Acemoglu2002}
%	\item \bibentry{Mortensen2003}
%	\item \bibentry{Mccall1970}
%	\item \bibentry{Rothschild1973}
%	\item \bibentry{Diamond1971}
%	\item \bibentry{Krueger2016}
%	\item \bibentry{Krueger2002}
%	\item \bibentry{Carling2001}
%	\item \bibentry{Chetty2008}
%	\item \bibentry{Card2015}	
%	\item \bibentry{Kolsrud2015}
%	\item \bibentry{Lalive2007}
%	\item \bibentry{VanOurs2008}	
%	\item \bibentry{Schmieder2016}
%	\item \bibentry{Nekoei2017}
%	\item \bibentry{Ljungqvist1998}	
%	
%	\item \bibentry{Diamond1981}
%	\item \bibentry{Diamond1982}
%	\item \bibentry{Diamond1982a}
%	\item \bibentry{Mortensen1982}
%	\item \bibentry{Mortensen1982a}
%	\item \bibentry{Pissarides1984}
%	\item \bibentry{Pissarides1985}
%	\item \bibentry{Petrongolo2001}
%	\item \bibentry{Gavazza2016}
%	\item \bibentry{Nash1950}
%	\item \bibentry{Hagedorn2016}
%	\item \bibentry{Chodorow-Riech2017}	
%	\item \bibentry{Hosios1990}
%	\item \bibentry{Hall2008}
%	\item \bibentry{Pissarides2009}
%	\item \bibentry{Haefke2013}
%	\item \bibentry{Kudlyak2014}
%	\item \bibentry{Merz1995}
%	\item \bibentry{Ravenna2011}
%	\item \bibentry{Mortensen1994}
%	\item \bibentry{Moen1997}
%	\item \bibentry{Menzio2011}
%	\item \bibentry{Sahin2014}
%
%	\item \bibentry{Jolivet2006}
%	\item \bibentry{Bontemps2000}
%	\item \bibentry{Mortensen2003}
%	\item \bibentry{Barlevy2008}
%	\item \bibentry{Engbom2017}
%	\item \bibentry{Sorkin2017}
%	\item \bibentry{Stevens2004}
%	\item \bibentry{Burdett2003}
%	\item \bibentry{Papp2013}
%	\item \bibentry{Williamson2010}
%	\item \bibentry{Lagos2009}
%	\item \bibentry{Shimer2000}
%\end{itemize}
%
%\bibliographystyle{apalike}
%\bibliography{bibtex}

\end{document}